Sequential Composition for Relaxed Memory: Pomsets with Predicate Transformers

### 2021-07-07 paper/intro.tex
cocktail time. only one todo left, and it is in an appendix
added some comments from mark and simon

To manage complexity, we have layered the definitions.  After an overview, we
define sequential dependencies in \textsection\ref{sec:model}.  The next two
sections add concurrency.  In \textsection\ref{sec:mca}, we define \PwTmca{},
which provides a Java-like model for \mca{} hardware, similar to that of
\citet{DBLP:journals/pacmpl/JagadeesanJR20}.  In \textsection\ref{sec:c11},
we define \PwTc{}, which models \cXI, adapting the approach of
\citet{DBLP:conf/esop/PaviottiCPWOB20}.  In \textsection\ref{sec:tool} we 
provide a tool for automatic evaluation of litmus tests.  In \textsection\ref{sec:additional},
we extend the semantics to include additional features, such as address
calculation.  In \textsection\ref{sec:results}, we summarize our results.

@@

For example, we can recover the expected semantics
for \eqref{eq1} by choosing
the predicate transformer which is independent of $(\DR x0)$
but dependent on $(\DR y0)$, which is the transformer
which maps $\bForm$ to $(0{=}s) \limplies \bForm$.
\begin{gather*}
  \PR{x}{r}\SEMI \PR{y}{s}\SEMI \IF{s{<}1} \THEN \PW{z}{r{*}s} \FI
  \\[-2ex]
  \nonumber
  \hbox{\begin{tikzinline}[node distance=1ex and 1.2em]
      \xform{rx0d}{(0{=}r) \limplies \bForm}{}
      \event{rx0}{\DR{x}{0}}{right=of rx0d}
      \xform{rx0ry0d}{(0{=}r) \limplies(0{=}s) \limplies \bForm}{right=of rx0}
      \event{ry0}{\DR{y}{0}}{right=of rx0ry0d}
      \xform{ry0d}{(0{=}s) \limplies \bForm}{right=of ry0}
      \event{wz0}{(0{=}s) \limplies (s {<} 1) \land (r{*}s){=}0 \mid \DW{z}{0}}{right=of ry0d}
      \xform{wz0i}{\bForm}{left=of rx0d}
      \po[out=20,in=170]{ry0}{wz0}
      \xo{rx0}{rx0d}
      \xo{ry0}{ry0d}
      \xo{rx0}{rx0ry0d}
      \xo{ry0}{rx0ry0d}
    \end{tikzinline}}
\end{gather*}
As a sanity check, we can see that sequential composition is
associative in this case, since it does not matter whether we
associate to the left, with intermediate step:
\begin{gather*}
  \PR{x}{r}\SEMI \PR{y}{s}
  \\
  \nonumber
  \hbox{\begin{tikzinline}[node distance=1ex and 1.5em]
      \xform{rx0d}{(0{=}r) \limplies \bForm}{}
      \event{rx0}{\DR{x}{0}}{right=of rx0d}
      \xform{rx0ry0d}{(0{=}r) \limplies(0{=}s) \limplies \bForm}{right=of rx0}
      \event{ry0}{\DR{y}{0}}{right=of rx0ry0d}
      \xform{ry0d}{(0{=}s) \limplies \bForm}{right=of ry0}
      \xform{wz0i}{\bForm}{left=of rx0d}
      \xo{rx0}{rx0d}
      \xo{ry0}{ry0d}
      \xo{rx0}{rx0ry0d}
      \xo{ry0}{rx0ry0d}
    \end{tikzinline}}
\end{gather*}
or to the right, with intermediate step:
\begin{gather*}
  \PR{y}{s}\SEMI \IF{s{<}1} \THEN \PW{z}{r{*}s} \FI
  \\
  \nonumber
  \hbox{\begin{tikzinline}[node distance=1ex and 1.5em]
      \event{ry0}{\DR{y}{0}}{}
      \event{wz0}{(0{=}s) \limplies (s {<} 1) \land (r{*}s){=}0 \mid \DW{z}{0}}{right=of ry0}
      \xform{ry0d}{(0{=}s) \limplies \bForm}{left=of ry0}
      \xform{wz0i}{\bForm}{left=of ry0d}
      \po{ry0}{wz0}
      \xo{ry0}{ry0d}
    \end{tikzinline}}
\end{gather*}
This is an instance of the general result that sequential composition forms a monoid.

\subsection{Related Work}

>>>>

To manage complexity, we have layered the definitions.  After an overview, we
define sequential dependencies in \textsection\ref{sec:model}.  The next two
sections add concurrency.  In \textsection\ref{sec:mca}, we define \PwTmca{},
which provides a Java-like model for \mca{} hardware, similar to that of
\citet{DBLP:journals/pacmpl/JagadeesanJR20}.  In
\textsection\ref{sec:results}, we summarize the results for this model. In \textsection\ref{sec:c11},
we define \PwTc{}, which models \cXI, adapting the approach of
\citet{DBLP:conf/esop/PaviottiCPWOB20}.  In \textsection\ref{sec:tool} we 
provide a tool for automatic evaluation of litmus tests.  In \textsection\ref{sec:additional},
we extend the semantics to include additional features, such as address
calculation.

@@

For example, we can recover the expected semantics
for \eqref{eq1} by choosing
the predicate transformer which is independent of $(\DR x0)$
but dependent on $(\DR y0)$, which is the transformer
which maps $\bForm$ to $(0{=}s) \limplies \bForm$.
\begin{gather*}
  \PR{x}{r}\SEMI \PR{y}{s}\SEMI \IF{s{<}1} \THEN \PW{z}{r{*}s} \FI
  \\
  \nonumber
  \hbox{\begin{tikzinline}[node distance=2ex and 1.2em]
      \xform{wz0i}{\bForm}{}
      \xform{rx0d}{(0{=}r) \limplies \bForm}{right=of wz0i}
      \xform{rx0ry0d}{(0{=}r) \limplies(0{=}s) \limplies \bForm}{right=of rx0d}
      \xform{ry0d}{(0{=}s) \limplies \bForm}{right=of rx0ry0d}
      \event{rx0}{\DR{x}{0}}{above left=of rx0ry0d}
      \event{ry0}{\DR{y}{0}}{above right=of rx0ry0d}
      \event{wz0}{(0{=}s) \limplies (s {<} 1) \land (r{*}s){=}0 \mid \DW{z}{0}}{right=of ry0}
      \po{ry0}{wz0}
      \xo{rx0}{rx0d}
      \xo{ry0}{ry0d}
      \xo{rx0}{rx0ry0d}
      \xo{ry0}{rx0ry0d}
    \end{tikzinline}}
\end{gather*}
In the diagram, the dotted lines indicate set inclusion into the index of the transformer-family.
As a sanity check, we can see that sequential composition is
associative in this case, since it does not matter whether we
associate to the left, with intermediate step:
\begin{gather*}
  \PR{x}{r}\SEMI \PR{y}{s}
  \\
  \nonumber
  \hbox{\begin{tikzinline}[node distance=2ex and 1.5em]
      \xform{rx0d}{(0{=}r) \limplies \bForm}{}
      \event{rx0}{\DR{x}{0}}{right=of rx0d}
      \xform{rx0ry0d}{(0{=}r) \limplies(0{=}s) \limplies \bForm}{right=of rx0}
      \event{ry0}{\DR{y}{0}}{right=of rx0ry0d}
      \xform{ry0d}{(0{=}s) \limplies \bForm}{right=of ry0}
      \xform{wz0i}{\bForm}{left=of rx0d}
      \xo{rx0}{rx0d}
      \xo{ry0}{ry0d}
      \xo{rx0}{rx0ry0d}
      \xo{ry0}{rx0ry0d}
    \end{tikzinline}}
\end{gather*}
or to the right, with intermediate step:
\begin{gather*}
  \PR{y}{s}\SEMI \IF{s{<}1} \THEN \PW{z}{r{*}s} \FI
  \\
  \nonumber
  \hbox{\begin{tikzinline}[node distance=1ex and 1.5em]
      \event{ry0}{\DR{y}{0}}{}
      \event{wz0}{(0{=}s) \limplies (s {<} 1) \land (r{*}s){=}0 \mid \DW{z}{0}}{right=of ry0}
      \xform{ry0d}{(0{=}s) \limplies \bForm}{left=of ry0}
      \xform{wz0i}{\bForm}{left=of ry0d}
      \po{ry0}{wz0}
      \xo{ry0}{ry0d}
    \end{tikzinline}}
\end{gather*}
This is an instance of the general result that sequential composition forms a monoid.

\section{Related Work}
